---
title: "Production sources year coverage"
output: rmarkdown::html_vignette
vignette: >
  %\VignetteIndexEntry{Production sources year coverage}
  %\VignetteEngine{knitr::rmarkdown}
  %\VignetteEncoding{UTF-8}
---

```{r, include = FALSE}
knitr::opts_chunk$set(
  collapse = TRUE,
  comment = "#>"
)
```

```{r setup}
library(PackageExampleR)
key_path <- here::here(Sys.getenv("GOOGLESHEETS_AUTH_FILE"))
googlesheets4::gs4_auth(path = key_path)
```

First we read the production sources sheet and build a dataframe where each row accounts for one year.

```{r}
# Step 1: Authentication
sheet_url <- "1UdwgS87x5OsLjNuKaY3JA01GoI5nwsenz62JXCeq0GQ"

# PART 2: production_sources FOR PRODUCTION
expanded_production_sources <-
  sheet_url |>
  googlesheets4::read_sheet(sheet = "Final_Sources_Production") |>
  expanded_production_sources()

expanded_production_sources
```

Now we build some plots.

Plot showing years covered by `expanded_production_sources`:

```{r, fig.alt="Plot showing years covered by expanded_production_sources"}
ggplot2::ggplot(
  expanded_production_sources,
  ggplot2::aes(y = Information_Level, x = Year, fill = "lightblue")
) +
  ggplot2::geom_tile(alpha = .8) +
  ggplot2::theme_dark() +
  ggplot2::labs(title = "Source Availability by Country") +
  ggplot2::scale_fill_identity() +
  ggplot2::facet_wrap(~Reporter, ncol = 1)
```

Plot showing by years, colored by `expanded_production_sources`:

```{r, fig.alt="Plot showing by years, colored by expanded_production_sources"}
ggplot2::ggplot(
  expanded_production_sources,
  ggplot2::aes(y = Access, x = Year, fill = Name, alpha = .8)
) +
  ggplot2::geom_tile() +
  # theme_minimal() +
  ggplot2::theme(legend.position = "None") +
  ggplot2::labs(title = "Source Availability by Country") +
  ggplot2::facet_grid(Info_Format ~ Reporter, scales = "free", space = "free_x")
```

Plot showing years covered by `expanded_production_sources`, colored by quality:

```{r, fig.alt="Plot showing years covered by expanded_production_sources, colored by quality"}
ggplot2::ggplot(
  expanded_production_sources, ggplot2::aes(y = Information_Level, x = Year, alpha = .8)
) +
  ggplot2::geom_tile(ggplot2::aes(fill = factor(Quality))) +
  ggplot2::theme_dark() +
  ggplot2::labs(title = "Source Availability by Country", fill = "Quality") +
  ggplot2::guides(alpha = "none") +
  ggplot2::scale_fill_manual(
    values = c("Bad" = "red", "Good" = "green", "Unknown" = "blue")
  ) +
  ggplot2::facet_grid(Info_Format ~ Reporter, scales = "free", space = "free_x")
```

Plot showing years covered by `expanded_production_sources`, colored by scanned/not scanned:

```{r, fig.alt="Plot showing years covered by expanded_production_sources, colored by scanned/not scanned"}
ggplot2::ggplot(expanded_production_sources, ggplot2::aes(y = Information_Level, x = Year)) +
  ggplot2::geom_tile(ggplot2::aes(fill = factor(Scanned), alpha = .8)) +
  ggplot2::theme_dark() +
  ggplot2::labs(title = "Scanned sources_prodcution by Country", fill = "Scanned") +
  ggplot2::scale_fill_manual(
    labels = c("0" = "No", "1" = "Yes"), values = c("0" = "red", "1" = "green")
  ) +
  ggplot2::guides(alpha = "none") +
  ggplot2::facet_grid(Info_Format ~ Reporter, scales = "free", space = "free_x")
```

Plot showing years covered by `expanded_production_sources`, colored by whether or not in SACO:

```{r, fig.alt="Plot showing years covered by expanded_production_sources, colored by whether or not in SACO"}
ggplot2::ggplot(
  expanded_production_sources,
  ggplot2::aes(y = Information_Level, x = Year, alpha = .8)
) +
  ggplot2::geom_tile(ggplot2::aes(fill = factor(In_Saco))) +
  ggplot2::theme_dark() +
  ggplot2::labs(title = "Source Availability by Country", fill = "In Saco") +
  ggplot2::guides(alpha = "none") +
  ggplot2::scale_fill_manual(
    labels = c("0" = "No", "1" = "Yes"), values = c("0" = "red", "1" = "green")
  ) +
  ggplot2::facet_grid(Info_Format ~ Reporter, scales = "free", space = "free_x")
```
